Sparsity Weak Lensing $3$-D Density Map Reconstruction: Reducing the line of sight smearing.

### 2020-10-27 doc/paper_ms_method_HSCY1/ms.tex
update the figures in paper

\usepackage{textcomp}
\usepackage{epstopdf}
\usepackage{natbib}
\usepackage{commath}
\hypersetup{breaklinks}
\DeclareMathOperator{\arccosh}{arccosh}
\newcommand \redColor{\color{red}}
\renewcommand\labelenumi{(\roman{enumi})}
\renewcommand\theenumi\labelenumi

\usepackage{algorithm,algcompatible}

@@

@@

\begin{abstract}
A new method is developed to reconstruct $3$-D density contrast maps from photometric weak-lensing shear measurements.
The $3$-D density contrast maps are modeled as a summation of the NFW basis atoms which have $2$-D multi-scale NFW surface
density profiles on the transverse plane and $1$-D Dirac delta functions in the line of sight direction. With the prior
assumption that the density fields can be sparsely represented by the NFW basis atoms, it is reconstructed using an
oracle algorithm: adaptive lasso. Our method is tested with realistic simulations using the HSC-like shape noise and
photo-$z$ uncertainty.
Our findings are summarized as folllows: 1) The global minima solution of the standard lasso algorithm suffers from a
smear of structure in the line of sight direction even in the absence of shape nosie.
2) The global minima solution of addaptive lasso algorithm significantly removes the line of sight smear of structure.
\end{abstract}

@@

\begin{figure*}[!t]
    \includegraphics[width=1.\textwidth]{nfwlet-atom-2D.pdf}
    \caption{The smoothed basis atoms. The first row shows the smoothed basis atoms in Fourier space
            and the second row shows the smoothed basis atoms in Real space.
            The first column is point mass atom and the other columns are multi-scale NFW atoms.
            The smoothing kernel is Gaussian with scale of $1.5$ pixels.} \label{fig-atoms2D}
\end{figure*}

\begin{figure}
 \includegraphics[width=0.5\textwidth]{nfwlet-atom-1D.pdf}
 \caption{The $1$-D slices for smoothed basis atoms at $x=0$. The corresponding $2$-D profiles are shown in Figure
        \ref{fig-atoms2D}.} \label{fig-atoms1D}
\end{figure}

@@

The primary goal of most weak-lensing surveys is to constrain the cosmology model through $2$-point correlations. The studies
include galaxy-galaxy lensing which cross correlating the shear field ($\gamma$) with the positions of foreground galaxies
\citep{gglens-GAMA-Han2014,gglens-BossCFHTMore2015,gglens-DES1},
and cosmic shear which auto-correlates the shear measurements
\citep{cosmicShearRealKids450,cosmicShear-DES1,cosmicShear_HSC1_Chiaki2019,cosmicShear_HSC1_Hamana2019}.
Since shear is directly related to the matter distribution as shown in eq. (\ref{eq-intro-delta2shear}), Galaxy-galaxy lensing
probes into the correlation between the matter field and galaxy field, on the other hand, cosmic shear probes into the
auto-correlation of matter field.

@@

In order to fully reconstruct the $3$-D mass density distribution ($\delta$) from the photometric shear observations ($\gamma$),
the density contrast field is modeled as a summation of basis atoms in a model dictionary
\begin{equation} \label{eq-intro-dict}
 \delta= \mathbf{\Phi} x,
\end{equation}
where $\mathbf{\Phi}$ is the transformation operator from the parameters in the dictionary space to the density contrast
and $x$ denotes the parameters. \citet{LSS-massMap-Wiener-Simon2009} reconstruct the density field
in Fourier space, which is equivalent to model the mass field with sinusoidial functions. On the other hand,
\citet{LSS-massMap-Glimpse3D-Leonard2014} models the mass field with Starlets \citep{Starlet-Starck2015}.

@@

\begin{figure}[!t]
 \centering
 \includegraphics[width=0.5\textwidth]{photo-z_binning.pdf}
 \caption{The source galaxies are binned into $10$ redshift bins according to their MLZ best photo-$z$ estimation.
        The blue histogram is the number distribution of the best photo-$z$ estimation. The vertical dashed lines
        are the bounds of bins. The galaxies are evenly distributed in each bins.} \label{fig-bestpz}
\end{figure}

@@

\begin{figure*}[!t]
\centering
\subfigure[lasso]{\includegraphics[width=0.49\textwidth]{delta-3-7-pz-nn-lasso.pdf}}
\subfigure[adaptive lasso]{\includegraphics[width=0.49\textwidth]{delta-3-7-pz-nn-alasso.pdf}}
\caption{The result of density map reconstruction with standard lasso (left) and adaptive lasso (right). The input density
        map is from a NFW halo with mass $M_{200}=10^{15} ~h^{-1}M_{\odot}$ at redshift $0.35$. The lasso reconstruction
        smear the density field along the line of sight direction (vertical direction in the plots). Noises on shear measurements
        are neglected in this simulation.} \label{fig-lassoVsadaLasso}
\end{figure*}

@@

Figure \ref{fig-bestpz} shows the histogram of the best MLZ photo-$z$ estimation \cite{HSC1-photoz} for galaxies in tract
9347 of HSC S16A data release \citep{HSC1-data}. These Galaxies are divided into ten source galaxy bins according to the
photo-$z$ best estimation and the boundaries of the bins are shown as vertical dash lines in Figure \ref{fig-bestpz}.
Figure \ref{fig-pdfpz} shows the average probability density function (PDF) in for galaxies in each redshift bin.

@@

\begin{figure}[!t]
 \centering
 \includegraphics[width=0.45\textwidth]{shapeMeasurementError-HSCY1.pdf}
 \caption{The Histogram of HSC-like shear measurement error (including errors from shape noise and photon noise) on the
        first component of shear ($g_1$).}
 \label{fig:mass-redshift}
\end{figure}

@@

\begin{figure}[!t]
 \centering
 \includegraphics[width=0.45\textwidth]{noise_std_map_pix.pdf}
 \caption{The standard deviation map of shear measurement error for the fifth source bin ($0.69 \leq z < 0.80 $).}
\end{figure}

@@

\begin{figure*}[!t]
 \centering
 \includegraphics[width=0.95\textwidth]{detfalseRate_f3-1.pdf}
 \includegraphics[width=0.95\textwidth]{detfalseRate_f3-3.pdf}
 \caption{The upper panels show the detection rate and false peak per square degree for mass map reconstruction with
NFW atoms and $\lambda=3.5$. The low panels show the results for $\lambda=5.0$. }
\end{figure*}

@@

\begin{figure*}[!t]
 \centering
 \includegraphics[width=0.95\textwidth]{peak_scatters_f3-1.pdf}
 \caption{The left panel shows the stacked distribution of deviations of detected peak positions
 from the centers of the corresponding input halos. The $x$-axis is for the deviated distance in the transverse plane and the
 $y$-axis is for the deviation of the redshift. The peaks inside the dashed black box are regarded as true detections. The
 right panel focus on the deviation of detected peaks in the line of sight direction. The $x$-axis shows the input redshifts
 and the $y$-axis is the redshift of the detected peak. The cross symbols show the average redshift of detected peak for
 halos with different input redshift and mass and the error-bars are the scatter of the corresponding peak redshifts. All
 of the results in this figure are based on the NFW atoms.}
\end{figure*}

@@

\begin{algorithm}[H]
\renewcommand{\thealgorithm}{}
\label{alg-1}
\caption{Our Algorithm}
\begin{algorithmic}[1]
\INPUT $\gamma$: Pixelized complex $3$-D array of shear
\OUTPUT  $\delta$: $3$-D array of density contrast
\STATE Normalize column vectors of $\mathbf{A}$
\STATE Estimate step size $\mu$ and $\mathbf{\Sigma}$
\STATE \textbf{Initialization:} ‎
\STATE $x'^{(1)} = 0$
\STATE $\hat{w}=1$, $\lambda=\lambda_{\rm{ls}}$
\STATE $t^{(1)}=1$,$i=1$, $j=1$
\WHILE{$j \leq 2$}
    \WHILE{$i \leq N_{\rm{iter}}$}
        \STATE $x'^{(n+1)}_{i}=\mathrm{ST}_{\hat{w}\lambda} \left(x'^{(n)}_{i} -\mu \partial_i G(x'^{(n)})\right)$
        \STATE $t^{(n+1)}=\frac{1+\sqrt{1+4(t^{(n)})^2}}{2}$
        \STATE $x'_{n+1} \leftarrow x'^{(x+1)}+ \frac{t^{(n)}-1}{t^{(n+1)}}(x'^{(n+1)}-x'^{(n)})$
        \STATE $i=i+1$
    \ENDWHILE
\STATE \textbf{Reinitialization:} ‎
\STATE $\hat{w}=\abs{\hat{x'}^{\rm{lasso}}}^{-2}$, $\lambda=\lambda_{\rm{als}}$
\STATE $\hat{x'}^{(1)} = x'^{(N_{\rm{iter}})}$
\STATE $t^{(1)}=1$, $i=1$
\STATE $j=j+1$
\ENDWHILE
\STATE $\delta=\mathbf{\Phi}\mathcal{N}^{-\frac{1}{2}}x'^{(N_{\rm{iter}})}$
\end{algorithmic}
\end{algorithm}

@@

\begin{figure*}[!ht]
 \centering
 \includegraphics[width=0.95\textwidth]{detfalseRate_f1-1.pdf}
 \includegraphics[width=0.95\textwidth]{detfalseRate_f1-3.pdf}
 \caption{The upper panels show the detection rate and false peak per square degree for mass map reconstruction with
point mass atoms and $\lambda=3.5$. The low panels show the results for $\lambda=5.0$. }
\end{figure*}

@@

The galay catalogs are simulated with the HSC S16A shape catalog \citep{HSC1-catalog}.
We uses galaxies in a one square degree region at the center of tract 9347 \citep{HSC1-data}.
By randomly rotating the galaxies in the shape catalog, we simulate HSC-like shear estimation errors
with different realizations.
The positions of the galaxies are randomized so that the galaxies homogeneously distributed in a one square degree
stamp.
For each galaxy, we randomly assign its redshift following the MLZ photo-$z$ probability distribution function
\citep{HSC1-photoz} of the galaxy.

@@

\begin{figure*}[!t]
\centering
\subfigure[Point Mass: $\lambda=3.5$]{\includegraphics[width=0.45\textwidth]{delta-1-7-pz-wn-PM-falsepeakproblem.pdf}}
\subfigure[NFW: $\lambda=3.5$]{\includegraphics[width=0.45\textwidth]{delta-1-7-pz-wn-NFW-falsepeakproblem.pdf}}
\subfigure[Point Mass: $\lambda=5.0$]{\includegraphics[width=0.45\textwidth]{delta-1-7-pz-wn-PM_3-falsepeakproblem.pdf}}
\subfigure[NFW: $\lambda=5.0$]{\includegraphics[width=0.45\textwidth]{delta-1-7-pz-wn-NFW_3-falsepeakproblem.pdf}}
\caption{The result of density map reconstruction with  Point Mass atoms (left) and  NFW atoms (right). The input density
        map is from a NFW halo with mass $M_{200}=10^{15.02} ~h^{-1}M_{\odot}$ at redshift $0.164$.
        } \label{fig-lassoVsadaLasso}
\end{figure*}

>>>>

\usepackage{textcomp}
\usepackage{epstopdf}
\usepackage{natbib}
\usepackage{commath}
\hypersetup{breaklinks}

\DeclareMathOperator{\arccosh}{arccosh}
\newcommand \redColor{\color{red}}
\renewcommand\labelenumi{(\roman{enumi})}
\renewcommand\theenumi\labelenumi

\usepackage{algorithm,algcompatible}
\usepackage[-]{callouts}

@@

\newcommand{\includezraphics}[1]{
\begin{annotate}{
\includegraphics[width=0.43\textwidth]{#1}}{0.45}
\arrow { -6., -4}{ -6., 4}
\note {-6.,4.5}{$\huge{z}$}
\end{annotate}
}

@@

\begin{abstract}
A new method is developed to reconstruct $3$-D density contrast maps from photometric weak-lensing shear measurements.
The $3$-D density contrast maps are modeled as a summation of the NFW basis atoms which have $2$-D multi-scale NFW surface
density profiles on the transverse plane and $1$-D Dirac delta functions in the line of sight direction. With the prior
assumption that the density fields can be sparsely represented by the NFW basis atoms, the density fields are reconstructed 
using an oracle algorithm: adaptive lasso. Our method is tested with realistic simulations using the HSC-like shape noise and
photo-$z$ uncertainty.
Our findings are summarized as folllows: 1) The global minima solution of the standard lasso algorithm suffers from a
smear of structure in the line of sight direction even in the absence of shape nosie.
2) The global minima solution of addaptive lasso algorithm significantly removes the line of sight smear of structure.
\end{abstract}

@@

\begin{figure*}[!t]
    \includegraphics[width=1.\textwidth]{nfwlet-atom-2D.pdf}
    \caption{The smoothed pixelized basis atoms. The upper row shows the basis atoms in Fourier space
            and the lower row shows the basis atoms in Real space.
            The leftmost column is point mass atom and the other columns are multi-scale NFW atoms.
            The smoothing kernel is Gaussian with standard deviation of $1.5$ pixels.} \label{fig-atoms2D}
\end{figure*}

\begin{figure}
 \includegraphics[width=0.5\textwidth]{nfwlet-atom-1D.pdf}
 \caption{The $1$-D slices for smoothed pixelized basis atoms at $x=0$. The corresponding $2$-D profiles are 
        shown in Figure \ref{fig-atoms2D}.} \label{fig-atoms1D}
\end{figure}

@@

The primary goal of most weak-lensing surveys is to constrain the cosmology model through $2$-point correlations. The studies
include galaxy$-$galaxy lensing which cross correlating the shear field ($\gamma$) with the positions of foreground galaxies
\citep{gglens-GAMA-Han2014,gglens-BossCFHTMore2015,gglens-DES1},
and cosmic shear which auto-correlates the shear measurements
\citep{cosmicShearRealKids450,cosmicShear-DES1,cosmicShear_HSC1_Chiaki2019,cosmicShear_HSC1_Hamana2019}.
Since shear is directly related to the matter distribution as shown in eq. (\ref{eq-intro-delta2shear}), Galaxy$-$galaxy lensing
probes into the correlation between the matter field and galaxy field, on the other hand, cosmic shear probes into the
auto-correlation of matter field.

@@

In order to fully reconstruct the $3$-D mass density distribution ($\delta$) from the photometric shear observations ($\gamma$),
the density contrast field is modeled as a summation of basis atoms in a model dictionary
\begin{equation} \label{eq-intro-dict}
 \delta= \mathbf{\Phi} x,
\end{equation}
where $\mathbf{\Phi}$ is the transformation operator from the projection parameters to the density contrast,
and $x$ denotes the parameters. \citet{LSS-massMap-Wiener-Simon2009} reconstruct the density field
in Fourier space, which is equivalent to model the mass field with sinusoidial functions. On the other hand,
\citet{LSS-massMap-Glimpse3D-Leonard2014} models the mass field with Starlets \citep{Starlet-Starck2015}.

@@

\begin{figure}[!t]
 \centering
 \includegraphics[width=0.5\textwidth]{photo-z_binning.pdf}
 \caption{The source galaxies are binned into $10$ redshift bins according to their MLZ best photo-$z$ estimation.
        The blue histogram is the number distribution of the best photo-$z$ estimation. The vertical dashed lines
        are the boundaries of the bins. The galaxies are equal-number binned.} \label{fig-bestpz}
\end{figure}

@@

\begin{figure*}[!t]
\centering
\subfigure[lasso]{\includezraphics{delta-3-7-pz-nn-lasso.pdf}}
\subfigure[adaptive lasso]{\includezraphics{delta-3-7-pz-nn-alasso.pdf}}
\caption{The result of density map reconstruction with standard lasso (left) and adaptive lasso (right). The input density
        map is from a NFW halo with mass $M_{200}=10^{15} ~h^{-1}M_{\odot}$ at redshift $0.35$. The lasso reconstruction
        smear the density field along the line of sight direction (vertical direction in the plots). Noises on shear measurements
        are neglected in this simulation.} \label{fig-lassoVsadaLasso}
\end{figure*}

@@

Figure \ref{fig-bestpz} shows the histogram of the best Machine Learning and photo-Z \citep[MLZ]{MLZ-TPZ2013} photometric redshift estimation 
from \cite{HSC1-photoz} for galaxies in the tract 9347 of the HSC S16A data release \citep{HSC1-data}. These Galaxies are divided 
into ten source galaxy bins according to the photo-$z$ best estimation and the boundaries of the bins are shown as vertical dash 
lines in Figure \ref{fig-bestpz}. Figure \ref{fig-pdfpz} shows the average probability density function (PDF) in for galaxies 
in each redshift bin.

@@

\begin{figure}[!t]
 \centering
 \includegraphics[width=0.45\textwidth]{shapeMeasurementError-HSCY1.pdf}
 \caption{The histograms of the HSC-like shear measurement error (including both from shape noise and from photon noise) 
        on the first component of shear ($g_1$) for galaxies (blue lines) and smoothed pixels (orange lines). The solid
        steps are the sampled histograms, and the dashed lines are the normalized gaussian distributions which have the
        same standard deviations as the corresponding histograms.}
 \label{fig:mass-redshift}
\end{figure}

@@

\begin{figure}[!t]
 \centering
 \includegraphics[width=0.45\textwidth]{noise_std_map_pix.pdf}
 \caption{The standard deviation pixel map of the HSC-like shear measurement error for the fifth source galaxy bin 
        ($0.69 \leq z < 0.80 $).}
\end{figure}

@@

\begin{figure*}[!t]
 \centering
 \includegraphics[width=0.95\textwidth]{detfalseRate_f3-1.pdf}
 \includegraphics[width=0.95\textwidth]{detfalseRate_f3-3.pdf}
 \caption{The upper panels show the halo detection rate and the number of false peaks per square degree for mass map 
 reconstruction with NFW atoms and $\lambda=3.5$. The low panels show the results for $\lambda=5.0$. }
\end{figure*}

@@

\begin{figure*}[!t]
 \centering
 \includegraphics[width=0.95\textwidth]{peak_scatters_f3-1.pdf}
 \caption{The left panel shows the stacked $2$-D distribution of the deviations of detected peak positions
 from the centers of the corresponding input halos. The $x$-axis is for the deviated distance in the transverse plane, and 
 the $y$-axis is for the deviation of the redshift. For each simulatin, the peaks inside the dashed black box with the
 minimal deviations (in unit of pixel) are defined as true detections. 
 The right panel focuses on the deviation of detected peaks in the line of sight direction. The $x$-axis is the input 
 halo redshifts, and the $y$-axis is the redshift of the detected peak. The `$\cross$' denotes the average redshift of 
 detected peaks for each halo over different realizations, and the error-bars are the scatters of the corresponding peak 
 redshifts. All of the results in this figure are based on the NFW dictionary with $\lambda=3.5$.}
\end{figure*}

@@

\begin{algorithm}[H]
\renewcommand{\thealgorithm}{}
\label{alg-1}
\caption{Our Algorithm}
\begin{algorithmic}[1]
\INPUT $\gamma$: Pixelized complex $3$-D array of shear
\OUTPUT  $\delta$: $3$-D array of density contrast
\STATE Normalize column vectors of $\mathbf{A}$
\STATE Estimate step size $\mu$ and $\mathbf{\Sigma}$
\STATE \textbf{Initialization:}
\STATE $x'^{(1)} = 0$
\STATE $\hat{w}=1$, $\lambda=\lambda_{\rm{ls}}$
\STATE $t^{(1)}=1$,$i=1$, $j=1$
\WHILE{$j \leq 2$}
    \WHILE{$i \leq N_{\rm{iter}}$}
        \STATE $x'^{(n+1)}_{i}=\mathrm{ST}_{\hat{w}\lambda} \left(x'^{(n)}_{i} -\mu \partial_i G(x'^{(n)})\right)$
        \STATE $t^{(n+1)}=\frac{1+\sqrt{1+4(t^{(n)})^2}}{2}$
        \STATE $x'_{n+1} \leftarrow x'^{(x+1)}+ \frac{t^{(n)}-1}{t^{(n+1)}}(x'^{(n+1)}-x'^{(n)})$
        \STATE $i=i+1$
    \ENDWHILE
\STATE \textbf{Reinitialization:}
\STATE $\hat{w}=\abs{\hat{x'}^{\rm{lasso}}}^{-2}$, $\lambda=\lambda_{\rm{als}}$
\STATE $\hat{x'}^{(1)} = x'^{(N_{\rm{iter}})}$
\STATE $t^{(1)}=1$, $i=1$
\STATE $j=j+1$
\ENDWHILE
\STATE $\delta=\mathbf{\Phi}\mathcal{N}^{-\frac{1}{2}}x'^{(N_{\rm{iter}})}$
\end{algorithmic}
\end{algorithm}

@@

\begin{figure*}[!ht]
 \centering
 \includegraphics[width=0.95\textwidth]{detfalseRate_f1-1.pdf}
 \includegraphics[width=0.95\textwidth]{detfalseRate_f1-3.pdf}
 \caption{The upper panels show the halo detection rate and the number of false peak per square degree for mass map 
 reconstruction with point mass atoms and $\lambda=3.5$. The low panels show the results for $\lambda=5.0$. }
\end{figure*}

@@

The galay catalogs are simulated with the HSC S16A shape catalog \citep{HSC1-catalog}.
We uses galaxies in a one square degree region at the center of tract 9347 \citep{HSC1-data}.
By randomly rotating the galaxies in the shape catalog, we simulate HSC-like shear estimation errors
with different realizations.
The positions of the galaxies are randomized so that the galaxies homogeneously distributed in a one square degree
stamp.
For each galaxy, we randomly assign its redshift following the MLZ photo-$z$ probability distribution function from
\citep{HSC1-photoz}.

@@

\begin{figure*}[!t]
\centering
\subfigure[Point Mass: $\lambda=3.5$]{\includezraphics{delta-1-7-pz-wn-PM-falsepeakproblem.pdf}}
\subfigure[NFW: $\lambda=3.5$]{\includezraphics{delta-1-7-pz-wn-NFW-falsepeakproblem.pdf}}
\subfigure[Point Mass: $\lambda=5.0$]{\includezraphics{delta-1-7-pz-wn-PM_3-falsepeakproblem.pdf}}
\subfigure[NFW: $\lambda=5.0$]{\includezraphics{delta-1-7-pz-wn-NFW_3-falsepeakproblem.pdf}}
\caption{The results of density maps reconstruction with Point Mass atoms (left) and  NFW atoms (right) with 
        $\lambda=3.5$ (upper) and $\lambda=5.0$ (lower). The input density
        map is from a NFW halo with mass $M_{200}=10^{15.02} ~h^{-1}M_{\odot}$ at redshift $0.164$.
        } \label{fig-lassoVsadaLasso}
\end{figure*}
